---
title: "Oxidative Stress"
subtitle: "Interactive data exploration"
---

```{r}
#| output: false

source(here::here("src", "setup.R"), echo = FALSE)

supplementary_data <- load_supplementary_data()

OS_data <- load_pcr_data("OS")
```

::: {.callout-tip appearance="simple"}
Most of the tables and figures on this page are **interactive** !
:::

# Raw Data

The complete raw data (before modeling)

```{r}
OS_data$raw |> 
  left_join(supplementary_data$gene_data$OS, join_by(Gene)) |> 
  left_join(select(OS_data$models, -Mod), join_by(Stage, Layer, Gene)) |> 
  select(Mouse, Experiment, Stage, Layer, Gene, Condition, Pathway, Mean_Cq, DCq, Fold) |> 
  datatable(
    rownames = FALSE,
    class = 'cell-border stripe compact',
    filter = 'top',
    options = list(pageLength = 10, autoWidth = TRUE)
  ) |> 
  formatRound(c("Mean_Cq", "DCq", "Fold"), 4)
```

## Data Dictionary

What each variable name represents, and how they are labeled within plots.

```{r}
make_data_dict_reactable(OS_data$clean, configs$data$PCR$data_dict)
```


# Data Composition

The count/proportion of data points within each level of our predictors.

```{r}
(OS_data$clean
  |> left_join(supplementary_data$gene_data$OS, join_by(Gene))
  |> droplevels()
  |> tbl_summary(
    include = c(Stage, Layer, Pathway),
    by = Condition
  )
  |> modify_header(label = "**Variable**") 
  |> bold_labels()
)
```

## Genes per Pathway

Which genes were included in each of our pathways.

```{r}
OS_data$clean |> 
  left_join(supplementary_data$gene_data$OS |> select(Pathway, Gene), join_by(Gene)) |> 
  make_suburst_plot(
    layers = c("Pathway", "Gene"),
    colors = sunburst_pcr_colors
  )
```


# Supplementary Gene Data

Supplementary information on the included genes.

```{r}
(OS_data$clean
  |> distinct(Gene)
  |> left_join(supplementary_data$gene_data$OS, join_by(Gene))
  |> select(-Figure)
  |> mutate(NCBI_reference = str_glue("<a target=_blank href='https://www.ncbi.nlm.nih.gov/gene/?term={NCBI_reference}'>{NCBI_reference}</a>"))
  |> arrange(Pathway, Gene)
  |> rename(
    `Gene Name` = Gene_name,
    `Putative Function` = Putative_function,
    `NCBI Reference` = NCBI_reference
  )
  |> mutate(Primers = str_glue("Frw: {Forward_primer}</br>Rev: {Reverse_primer}"), .keep = "unused")
  |> datatable(
    rownames = FALSE,
    class = "cell-border stripe compact",
    extensions = c("RowGroup", "Responsive"),
    filter = "top",
    options = list(pageLength = 5, autoWidth = TRUE, rowGroup = list(dataSrc = 1)),
    escape = FALSE
  )
)
```


# Data Summary

Summary of the DCq and Fold values for every `Gene`, at each `Stage` and `Layer`.

```{r}
OS_data_summary <- (
  OS_data$clean
  |> left_join(select(OS_data$models, -Mod), join_by(Stage, Layer, Gene))
  |> summarise(
    .by = c(Stage, Layer, Gene, Condition),
    Mean = mean(DCq),
    DCq = list(DCq), 
    Fold = first(Fold),
    BoxPlots = NA
  )
  |> pivot_wider(names_from = "Condition", values_from = c("Mean", "DCq", "Fold"))
  |> select(Stage, Layer, Gene, `Mean(N)` = Mean_N, `Mean(IH)` = Mean_IH, BoxPlots, DCq_N, DCq_IH, Fold = Fold_IH)
)
  
reactable(
  OS_data_summary |> select(-matches("DCq"))
  , defaultColDef = colDef(sortNALast = TRUE, vAlign = "center", headerVAlign = "center", align = "center")
  , columns = list(
    Stage = colDef(maxWidth = 75)
    , Layer = colDef(maxWidth = 75)
    , Gene = colDef(maxWidth = 100)
    , Condition = colDef(filterable = FALSE, maxWidth = 125)
    , `Mean(N)` = colDef(format = colFormat(digits = 2), filterable = FALSE, maxWidth = 100)
    , `Mean(IH)` = colDef(format = colFormat(digits = 2), filterable = FALSE, maxWidth = 100)
    , BoxPlots = colDef(
      cell = \(x, idx) {
        tags$div(
          sparkline(
            OS_data_summary$DCq_N[[idx]], type = "box", width = 525, tooltipFormatFieldlist = c('lw', 'lq', 'med', 'uq', 'rw'), 
            boxFillColor = "#02ccae", boxLineColor = "black", whiskerColor = "black", medianColor = "grey50", lineColor = "black",
            chartRangeMin = min(OS_data$clean$DCq), chartRangeMax = max(OS_data$clean$DCq)
          ),
          br(),
          sparkline(
            OS_data_summary$DCq_IH[[idx]], type = "box", width = 525, tooltipFormatFieldlist = c('lw', 'lq', 'med', 'uq', 'rw'),
            boxFillColor = "#9965db", boxLineColor = "black", whiskerColor = "black", medianColor = "grey50", lineColor = "black",
            chartRangeMin = min(OS_data$clean$DCq), chartRangeMax = max(OS_data$clean$DCq)
          ),
        )
      },
      sortable = FALSE, filterable = FALSE
    )
    , Fold = colDef(format = colFormat(digits = 2), filterable = FALSE, maxWidth = 75)
  )
  , showSortable = TRUE
  , filterable = TRUE
  , searchable = TRUE
  , showPageSizeOptions = TRUE
  , defaultPageSize = 10
  , minRows = 10
  , pageSizeOptions = c(10, 25, 50, 100)
  , outlined = TRUE
  , striped = TRUE
  , highlight = TRUE
  , compact = TRUE
  , fullWidth = TRUE
)
```


## Genes per Layer & Stage

Which genes were included in our pathways.

```{r}
OS_data$models |> 
  select(-Mod) |> 
  make_suburst_plot(
    layers = c("Layer", "Stage", "Gene"), 
    tooltips = c("Gene", "Fold"),
    colors = sunburst_pcr_colors,
    plot_options = list(insidetextorientation = 'radial')
  )
```


# DCq Distribution

The distribution of `DCq` values for a `Gene`, at a certain `Stage` and `Layer`

```{r}
OS_shared <- SharedData$new(OS_data$clean) # TODO: mutate(OS_data, ID = str_glue("Layer: {Layer}, Stage: {Stage}, Gene: {Gene}"), .by = c(Stage, Layer, Gene))

bscols(
  widths = c(1, 3, 3, 3),
  div(),
  filter_select("stage_filter_OS", "Stages", OS_shared, ~ Stage, multiple = FALSE),
  filter_select("layer_filter_OS", "Layers", OS_shared, ~ Layer, multiple = FALSE),
  filter_select("gene_filter_OS", "Genes", OS_shared, ~ Gene, multiple = FALSE)
)
```

```{js}
//| output: false

function filter_default() {
  document.getElementById("stage_filter_OS").getElementsByClassName("selectized")[0].selectize.setValue("P8", false)
  document.getElementById("layer_filter_OS").getElementsByClassName("selectized")[0].selectize.setValue("PC", false) 
  document.getElementById("gene_filter_OS").getElementsByClassName("selectized")[0].selectize.setValue("Apex1", false)
}

window.onload = filter_default;
```

```{r}
#| fig-width: 10

(plot_ly(
    OS_shared,
    x = ~ interaction(Condition, Gene, Layer, Stage),
    y = ~ DCq,
    color = ~ Condition,
    name = ~ Layer,
    split = ~ Stage,
    type = "box",
    boxpoints = "all", 
    jitter = 0.3,
    colors = c(N = "#02ccae", IH = "#9965db")
  )
  |> layout(
    autosize = TRUE, 
    showlegend = FALSE, 
    xaxis = list(title = NA, type = "category")
  )
)
```


# Temporal & Spatial Dynamics

```{r}
#| fig-width: 10
#| fig-height: 15

left_join(OS_data$clean, select(OS_data$models, -Mod), join_by(Stage, Layer, Gene)) |> 
  make_heatmap("Layer", "Stage")
```
