```{r}
source("../../src/setup.R", echo = FALSE)
```

<!---------------------------------------------------------->
<!---------------------------------------------------------->
# I. Data
***

```{r}
(supplementary_data <- load_supplementary_data())
```

```{r}
(ND_data <- load_pcr_data("ND", reprocess = TRUE, max_cq_clean = 33, refit = FALSE))
```


<!---------------------------------------------------------->
<!---------------------------------------------------------->
# II. Model fitting
***

Let's define the model we will fit to each `Gene`'s `DCq` data. 

Here, we will fit a simple Linear Model, which is largely similar to running a t-test between both conditions:

```{r}
ND_model <- function(data) {
  glmmTMB(DCq ~ Condition, family = gaussian("identity"), data = data, contrasts = list(Condition = "contr.sum"))
}
```

Now, let's fit said model to each `Gene`'s data, for a given `Stage` and `Layer`:

```{r}
compute_fold_change <- function(mod) {
  return(
    get_data(mod) 
    |> select(Condition, DCq) 
    |> pivot_wider(names_from = Condition, values_from = DCq, values_fn = mean) 
    |> summarize(Fold = 2**(-1 * (IH - N))) 
    |> pull(Fold) 
  )
}
```

```{r}
(ND_data$models <- ND_data$clean
  |> group_split(Stage, Layer, Gene)
  |> map_dfr(
    \(d) summarize(d, Mod = pick(Condition, DCq) |> ND_model() |> list(), .by = c(Stage, Layer, Gene)), 
    .progress = "Fitting models:"
  )
  |> filter(!has_na_coefs(Mod)) # Removing models that did not fit properly
  |> mutate(Fold = map_dbl(Mod, compute_fold_change)) # Adding the Fold change
  |> select(Stage, Layer, Gene, Fold, Mod)
)
```


<!---------------------------------------------------------->
<!---------------------------------------------------------->
# III. Model analysis
***

## Predictions

For each model we fit, we can then extract the CI and p.value for the relevant contrasts, and use those to establish if a `Gene` was up or down-regulated:

```{r}
get_emmeans_data <- function(mod) {
  return(
    emmeans(mod, specs = "Condition", type = "response")
    |> contrast(method = "pairwise", adjust = "none", infer = TRUE)
    |> as.data.frame()
    |> pivot_wider(names_from = contrast, values_from = estimate)
    |> select(last_col(), LCB = lower.CL, UCB = upper.CL, p.value)
    |> mutate(across(where(is.character), \(x) na_if(x, "NaN")))
  )
}
```

```{r}
(ND_data$predictions <- ND_data$models
  |> group_split(Stage, Layer, Gene)
  |> map_dfr(\(d) mutate(d, get_emmeans_data(Mod[[1]])), .progress = "Extracting model predictions:")
  |> filter(!is.na(p.value))
  |> mutate(Expression = get_regulation_type(Fold, p.value))
  |> select(Stage, Layer, Gene, Fold, Expression, matches("-|/"), LCB, UCB, p.value)
)
```

## Boxplots

Generating boxplots for every `Gene` (with a statistically significant difference between N and IH), split by `Layer` & `Stage`:

```{r fig.width = 15, fig.height = 20}
(
  ND_data$clean 
  |> left_join(ND_data$predictions, join_by(Stage, Layer, Gene))
  |> filter(p.value <= .05)
  |> group_by(Layer, Stage)
  |> group_map(\(d, g) {
    n.genes <- n_distinct(d$Gene)
    max.cols <- 6
    n.cols <- min(n.genes, max.cols)
    n.rows <- ceiling(n.genes / max.cols)
  
    make_boxplot_panel(d, n.cols, n.rows, display_labels = TRUE)
    }, .keep = TRUE
  )
)
```

## Gene regulation timeline

To get a better idea of how each `Gene`'s regulation changes through time, we can plot a timeline of their expression, split by `Layer` and `Pathway`.

```{r fig.width = 12, fig.height = 8}
(
  ND_data$predictions 
  |> left_join(supplementary_data$gene_data$ND, join_by(Gene))
  |> left_join(supplementary_data$layer_families, join_by(Layer))
  |> filter(p.value <= .05)
  |> filter(!(Layer_family == "PC" & Stage == "P4")) # PCs are too small and undifferentiated at that stage to properly microdissect
  |> select(Stage, Layer, Layer_family, Gene, Fold, p.value, Expression, Pathway, Pathway_family)
  |> mutate(Stage = case_when(
      Layer %in% c("EGLi", "EGLo") ~ str_glue("{Stage} ({Layer})"),
      .default = as.character(Stage)
    ) 
    |> factor(levels = c("P4", "P8", "P8 (EGLo)", "P8 (EGLi)", "P12", "P21", "P70"))
  )
  |> group_by(Layer_family)
  |> group_map(\(d, g) make_fold_timeline_plot(d, facet_rows = "Pathway_family", trans = "log", colors = colors_fold, size_boost = 1.5))
)
```

## Temporal & Spatial dynamics

```{r fig.width = 12, fig.height = 13}

left_join(ND_data$clean, ND_data$predictions, join_by(Stage, Layer, Gene)) |> 
  make_heatmap("Layer", "Stage")
```


## Regulation & Pathways

We can get a general overview of which `Gene` are up or down-regulated through a Sunburst plot, stacked by `Stage` and `Layer`, visualize to which `Pathway` the regulated genes are linked to.

```{r fig.width = 200, fig.height = 200}
ND_data$predictions |> 
  left_join(supplementary_data$gene_data$ND, join_by(Gene)) |>
  make_circlize_plot()
```
