```{r}
source("../../src/setup.R", echo = FALSE)
```

<!---------------------------------------------------------->
<!---------------------------------------------------------->
# I. Data
***

```{r}
(supplementary_data <- load_supplementary_data())
```

```{r}
(OS_data <- load_pcr_data("OS", reprocess = TRUE, max_cq_clean = 33, refit = FALSE))
```

<!---------------------------------------------------------->
<!---------------------------------------------------------->
# II. Model fitting
***

Let's define the model we will fit to each `Gene`'s `DCq` data. 

Here, we will fit a Linear Mixed Effect Model, with a random effect per Experiment, to account for potential clustering:

```{r}
OS_model <- function(data) {
  form <- "DCq ~ Condition"
  
  # If the data contains an Experiment variable with more than two levels, add a random effect by Experiment
  if ("Experiment" %in% colnames(data) && n_distinct(data$Experiment) >= 2) 
    form <- str_c(form, " + (1 | Experiment)")

  glmmTMB(as.formula(form), family = gaussian("identity"), data = data, contrasts = list(Condition = "contr.sum"))
}
```

Now, let's fit said model to each `Gene`'s data, for a given `Stage` and `Layer`:

```{r}
compute_fold_change <- function(mod) {
  return(
    get_data(mod)
    |> select(Condition, DCq)
    |> pivot_wider(names_from = Condition, values_from = DCq, values_fn = mean)
    |> summarize(Fold = 2**(-1 * (IH - N)))
    |> pull(Fold)
  )
}
```

```{r}
(OS_data$models <- OS_data$clean
  |> group_split(Stage, Layer, Gene)
  |> map_dfr(
    \(d) summarize(d, Mod = pick(Experiment, Condition, DCq) |> OS_model() |> list(), .by = c(Stage, Layer, Gene)), 
    .progress = "Fitting models:"
  )
  |> filter(!has_na_coefs(Mod)) # Removing models that did not fit properly
  |> mutate(Fold = map_dbl(Mod, compute_fold_change)) # Adding the Fold change
  |> select(Stage, Layer, Gene, Fold, Mod)
)
```

# Model analysis

## Predictions

For each model we fit, we can then extract the CI and p.value for the relevant contrasts, and use those to establish if a `Gene` was up or down-regulated:

```{r}
get_emmeans_data <- function(mod) {
  return(
    emmeans(mod, specs = "Condition", type = "response")
    |> contrast(method = "pairwise", adjust = "none", infer = TRUE)
    |> as.data.frame()
    |> pivot_wider(names_from = contrast, values_from = estimate)
    |> select(last_col(), LCB = lower.CL, UCB = upper.CL, p.value)
    |> mutate(across(where(is.character), \(x) na_if(x, "NaN")))
  )
}
```

```{r}
(OS_data$predictions <- OS_data$models
  |> group_split(Stage, Layer, Gene)
  |> map_dfr(\(d) mutate(d, get_emmeans_data(Mod[[1]])), .progress = "Extracting model predictions:")
  |> filter(!is.na(p.value))
  |> mutate(Expression = get_regulation_type(Fold, p.value))
  |> select(Stage, Layer, Gene, Fold, Expression, matches("-|/"), LCB, UCB, p.value)
)
```

## Boxplots

Generating boxplots for every `Gene` (with a statistically significant difference between N and IH), split by `Layer` & `Stage`:

```{r fig.width = 15, fig.height = 25}
(
  OS_data$clean 
  |> left_join(OS_data$predictions, join_by(Stage, Layer, Gene))
  |> filter(p.value <= .05)
  |> group_by(Layer, Stage)
  |> group_map(\(d, g) {
    n.genes <- n_distinct(d$Gene)
    max.cols <- 6
    n.cols <- min(n.genes, max.cols)
    n.rows <- ceiling(n.genes / max.cols)
  
    make_boxplot_panel(d, n.cols, n.rows, display_labels = TRUE)
    }, .keep = TRUE
  )
)
```


## Gene regulation timeline

To get a better idea of how each `Gene`'s regulation changes through time, we can plot a timeline of their expression, split by `Layer` and `Pathway`.

```{r fig.width = 12, fig.height = 8}
plot_OS_timeline <- function(dat, group, size_boost = 1.2) {
  
  if (group[[1]][1] == "Whole") {
    
    dat |>
      group_by(Figure) |> 
      group_map(\(d, g) plot_OS_timeline(d, str_glue("Whole Cerebellum - {first(g[[1]])}"), 1.5))
  }
  else {
    
    group_name <- group[[1]][1]
    if (group_name == "PC") group_name <- "Purkinje Cells"
    
    if (group_name == "Cellular Response") {
      cell_resp_levels <- c("Cell Death and Protection", "Apoptotic Pathways", "Inflammation Pathways", "Autophagy and Mitophagy")
      dat <- mutate(dat, Pathway = factor(Pathway, levels = cell_resp_levels))
    }
    
    make_fold_timeline_plot(dat, facet_rows = "Pathway", trans = "log", color_by = "Effect", size_boost = size_boost)
  }
}
```

### Whole Cerebellum

```{r fig.width = 12, fig.height = 8}
(
  OS_data$predictions 
  |> left_join(supplementary_data$gene_data$OS, join_by(Gene))
  |> left_join(supplementary_data$layer_families, join_by(Layer))
  |> filter(p.value <= .05 & Layer == "Whole")
  |> mutate(Effect = case_when(
      str_detect(Expression, "Downregulated") & Effect == "Beneficial" ~ "Deleterious",
      str_detect(Expression, "Downregulated") & Effect == "Deleterious" ~ "Beneficial",
      .default = Effect
    )
  )
  |> select(Stage, Layer, Gene, Fold, p.value, Expression, Effect, Pathway, Figure)
  |> group_by(Figure)
  |> group_map(\(d, g) {
      
      if (g$Figure[[1]] == "Cellular Response") 
        d <- mutate(d, Pathway = factor(Pathway, levels = c("Cell Death and Protection", "Apoptotic Pathways", "Inflammation Pathways", "Autophagy and Mitophagy")))
    
      make_fold_timeline_plot(d, facet_rows = "Pathway", trans = "log", color_by = "Effect", colors = colors_effect, size_boost = 2) +
        labs(title = g$Figure[[1]])
    }
  )
)
```

### Purkinje Cells

```{r fig.width = 9, fig.height = 6}
(
  OS_data$predictions 
  |> left_join(supplementary_data$gene_data$OS, join_by(Gene))
  |> left_join(supplementary_data$layer_families, join_by(Layer))
  |> filter(p.value <= .05 & Layer == "PC")
  |> mutate(Effect = case_when(
      str_detect(Expression, "Downregulated") & Effect == "Beneficial" ~ "Deleterious",
      str_detect(Expression, "Downregulated") & Effect == "Deleterious" ~ "Beneficial",
      .default = Effect
    )
  )
  |> select(Stage, Layer, Gene, Fold, p.value, Expression, Effect, Pathway)
  |> group_by(Layer)
  |> group_map(\(d, g) make_fold_timeline_plot(d, facet_rows = "Pathway", trans = "log", color_by = "Effect", size_boost = 1.5) + labs(title = g$Layer[[1]]))
)
```

# Temporal & Spatial dynamics

```{r fig.width = 8, fig.height = 15}

left_join(OS_data$clean, OS_data$predictions, join_by(Stage, Layer, Gene)) |> 
  make_heatmap("Layer", "Stage")
```
