```{r}
source("../../src/setup.R", echo = FALSE)

OS_fig_path <- "PCR/OS"
```

<!---------------------------------------------------------->
<!---------------------------------------------------------->
# I. Whole Cerebellum:
***

<!---------------------------------------------------------->
## 1. Data:
***

```{r}
(OS_Tot_data_raw <- load_OS_Tot(layer = "Tot", max_cq = 33, hk_genes = c("Hsp90", "Ppia"), update_with_cal = TRUE))
```

<!---------------------------------------------------------->
## 2. Models:
***

**Fit the chosen model to each Gene:**

```{r}
OS_Tot_model_cond <- function(data) {
  form <- "DCq ~ Condition"
  if ("Experiment" %in% colnames(data) && n_distinct(data$Experiment) >= 2) form <- str_c(form, " + (1 | Experiment)")

  return(glmmTMB(as.formula(form), family = gaussian("identity"), data = data, contrasts = list(Condition = "contr.sum")))
}

(OS_Tot_mods_cond <- OS_Tot_data_raw 
  |> filter(n() >= 3, .by = c(Stage, Layer, Gene, Condition)) # Checking that there are at least 3 values per Condition
  |> group_by(Stage, Layer, Gene)
  |> filter(n_distinct(Condition) >= 2) # Checking that both conditions are present for each Gene
  |> group_split()
  |> map_dfr(\(d) suppressMessages({summarize(d, mod = list(OS_Tot_model_cond(pick(everything()))), .by = c(Stage, Layer, Gene))}), .progress = "Fitting models:")
  |> filter(!has_na_coefs(mod))
  |> ungroup()
)
```

**Extract the contrasts, p.values, and CI:**

```{r}
(OS_Tot_data_cond <- OS_Tot_mods_cond
 |> group_split(Stage, Layer, Gene)
 |> map_dfr(\(d)
      mutate(d,
        emmeans(mod[[1]], specs = "Condition", type = "response") |> 
          contrast(method = "pairwise", adjust = "none", infer = TRUE) |> 
          as.data.frame() |> 
          select(Condition.cont = matches("estimate|risk|odds|^ratio|^difference"), Condition.LCB = lower.CL, Condition.UCB = upper.CL, Condition.p = p.value)
      ) |> select(-mod),
      .progress = "Extracting model predictions:"
  )
 |> mutate(across(where(is.character), \(x) na_if(x, "NaN")))
 |> right_join(OS_Tot_data_raw, by = join_by(Stage, Layer, Gene))
 |> filter(if_all(matches(".p$"), \(x) !is.na(x)))
 |> add_expression("Condition.p")
 |> mutate(Effect = case_when(
      str_detect(Expression, "Downregulated") & Effect == "Good" ~ "Bad",
      str_detect(Expression, "Downregulated") & Effect == "Bad" ~ "Good",
      .default = Effect
    )
  )
)
```

<!---------------------------------------------------------->
## 3. Plots:
***

**Generate boxplots (with p-values) for every gene, split by Layer & Stage:**

```{r}
OS_Tot_data_cond |> 
  group_by(Layer, Stage) |> 
  group_walk(\(d, g) {
    n.genes <- n_distinct(d$Gene)
    max.cols <- 10
    n.cols <- min(n.genes, max.cols)
    n.rows <- ceiling(n.genes / max.cols)
    
    make_boxplot_panel(d, n.cols, n.rows, display_labels = TRUE) |>
    save_png(
      glue("[OS-Tot] BoxPlots [Condition] - {g$Layer} & {g$Stage}"),
      subfolder = OS_fig_path, width = 1 + n.cols * 2, height = 1 + n.rows * 4
    )
  }, .keep = TRUE
)
```

**Generate Fold change timelines for each Layer and Pathway:**

```{r fig.width = 7, fig.height = 5}
OS_Tot_data_cond |> 
  filter(any(Condition.p <= alpha), .by = c(Gene, Figure)) |> 
  mutate(Fold = if_else(Condition.p <= alpha, Fold, NA)) |> 
  group_by(Figure) |> 
  group_walk(
    \(d, g) {
      width <- 2 + n_distinct(d$Stage)
      height <- d |> group_by(Pathway) |> group_map(\(d, g) n_distinct(d$Gene) * 0.1 + 1) |> flatten_dbl() |> sum()
      
      if (g$Figure[[1]] == "Cellular Response") 
        d <- mutate(d, Pathway = factor(Pathway, levels = c("Cell Death and Protection", "Apoptotic Pathways", "Inflammation Pathways", "Autophagy and Mitophagy")))
      
      make_vertical_fold_timeline(d, trans = "log", color_by = "Effect", colors = colors_effect, title = g$Figure[[1]], size_boost = 2) |> 
        save_png(
          glue("[OS-Tot] Timeline of Fold Change by [Pathway] - {str_replace_all(g$Figure[[1]], '/', ', ')}"), 
          subfolder = OS_fig_path, width = width, height = height
        )
    }
  )
```


<!---------------------------------------------------------->
<!---------------------------------------------------------->
***
# II. Purkinje Cells:
***

<!---------------------------------------------------------->
## 1. Data:
***

```{r}
(OS_PC_data_raw <- load_OS_PC(layer = "PC", max_cq = 40))
```


<!---------------------------------------------------------->
## 2. Models:
***

**Fit the chosen model to each Gene:**

```{r}
OS_PC_model_cond <- function(data) {
  form <- "DCq ~ Condition"
  if ("Experiment" %in% colnames(data) && n_distinct(data$Experiment) >= 2) form <- str_c(form, " + (1 | Experiment)")

  return(glmmTMB(as.formula(form), family = gaussian("identity"), data = data, contrasts = list(Condition = "contr.sum")))
}

(OS_PC_mods_cond <- OS_PC_data_raw 
  |> filter(n() >= 3, .by = c(Stage, Layer, Gene, Condition)) # Checking that there are at least 3 values per Condition
  |> group_by(Stage, Layer, Gene)
  |> filter(n_distinct(Condition) >= 2) # Checking that both conditions are present for each Gene
  |> group_split()
  |> map_dfr(\(d) suppressMessages({summarize(d, mod = list(OS_PC_model_cond(pick(everything()))), .by = c(Stage, Layer, Gene))}), .progress = "Fitting models:")
  |> filter(!has_na_coefs(mod))
  |> ungroup()
)
```

**Extract the contrasts, p.values, and CI:**

```{r}
(OS_PC_data_cond <- OS_PC_mods_cond
 |> group_split(Stage, Layer, Gene)
 |> map_dfr(\(d) 
      mutate(d,
        emmeans(mod[[1]], specs = "Condition", type = "response") |> 
          contrast(method = "pairwise", adjust = "none", infer = TRUE) |> 
          as.data.frame() |> 
          select(Condition.cont = matches("estimate|risk|odds|^ratio|^difference"), Condition.LCB = lower.CL, Condition.UCB = upper.CL, Condition.p = p.value)
      ) |> select(-mod),
      .progress = "Extracting model predictions:"
  )
 |> mutate(across(where(is.character), \(x) na_if(x, "NaN")))
 |> right_join(OS_PC_data_raw, by = join_by(Stage, Layer, Gene))
 |> filter(if_all(matches(".p$"), \(x) !is.na(x)))
 |> add_expression("Condition.p")
 |> mutate(Effect = case_when(
      str_detect(Expression, "Downregulated") & Effect == "Good" ~ "Bad",
      str_detect(Expression, "Downregulated") & Effect == "Bad" ~ "Good",
      .default = Effect
    )
  )
)
```


<!---------------------------------------------------------->
## 3. Plots:
***

**Generate boxplots (with p-values) for every gene, split by Layer & Stage:**

```{r}
OS_PC_data_cond |> 
  group_by(Layer, Stage) |> 
  group_walk(\(d, g) {
    n.genes <- n_distinct(d$Gene)
    max.cols <- 6
    n.cols <- min(n.genes, max.cols)
    n.rows <- ceiling(n.genes / max.cols)
  
    make_boxplot_panel(d, n.cols, n.rows, display_labels = TRUE) |>
    save_png(
      glue("[OS-PC] BoxPlots [Condition] - {g$Layer} & {g$Stage}"),
      subfolder = OS_fig_path, width = 1 + n.cols * 2, height = 1 + n.rows * 4
    )
  }, .keep = TRUE
)
```

**Generate Fold change timelines for each Layer and Pathway:**

```{r fig.width = 7, fig.height = 5}
OS_PC_data_cond |> 
  filter(any(Condition.p <= alpha), .by = c(Gene, Layer)) |> 
  mutate(Fold = if_else(Condition.p <= alpha, Fold, NA)) |> 
  group_by(Layer) |> 
  group_walk(
    \(d, g) {
      width <- 2 + n_distinct(d$Stage)
      height <- d |> group_by(Pathway) |> group_map(\(d, g) n_distinct(d$Gene) * 0.1 + 1) |> flatten_dbl() |> sum()
      make_vertical_fold_timeline(d, trans = "log", color_by = "Effect", title = g$Layer[[1]], size_boost = 1.5) |>
        save_png(
          glue("[OS-PC] Timeline of Fold Change by [Pathway] - {g$Layer[[1]]}"),
          subfolder = OS_fig_path, width = width, height = height
        )
    }
  )
```
