```{r}
source("../../src/setup.R", echo = FALSE)

Calb_P8_fig_path <- "IHC/Calb-P8"
stage = "P8"
```

<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
***
# I. Data:
***

```{r}
(Calb_P8_data <- load_Calb())

### Variables:

Calb_P8_responses <- c("N_CC", "Dens_CC", "Vol_PC_per_cell", "Vol_PC_per_cell", "A_PCL_ML", "A_Cortex")
Calb_P8_predictors <- c("Condition", "Mouse")

N_CC_name <- "Number of Purkinje cell bodies  \n *(per 250x10<sup>3</sup> µm<sup>3</sup>)*"
A_PC_per_cell_name <- "Purkinje marked area *(10<sup>-4</sup> µm²)*  \n per Purkinje cell"
Vol_PC_per_cell_name <- "Purkinje marked volume *(10<sup>-4</sup> µm<sup>3</sup>)*  \n per Purkinje cell"
A_PCL_ML_name <- "Area of the PCL and ML *(μm²)*"
A_Cortex_name <- "Area of the Cortex *(μm²)*"

contrasts(Calb_P8_data$Condition) <- contr.sum

### Averaged data:

(Calb_P8_data_agg <- Calb_P8_data |> 
  summarize(across(any_of(Calb_P8_responses), \(x) mean(x, na.rm = TRUE)), .by = c(Mouse, Condition)) |> 
  mutate(
    ID = row_number(),
    dCondition = as.numeric(Condition == "IH"),
    .before = 1
  )
)
```


<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
***
# II. N_CC / Dens:
***

<!------------------------------------------------------------------------------>
## 1. Models
***

### Welch t-test:

On data averaged by Mouse.
Data is assumed heteroscedastic in relation to `Condition`.

```{r}
Calb_P8_mod_N_CC_agg <- afex::mixed(
  N_CC ~ Condition + (0 + dCondition | ID),
  data = Calb_P8_data_agg,
  control = lmerControl(check.nobs.vs.nlev = "ignore", check.nobs.vs.nRE = "ignore"),
  check_contrasts = FALSE,
  method = "S",
  type = 3,
  test_intercept = FALSE
)

parameters(Calb_P8_mod_N_CC_agg$full_model, effects = "fixed", ci_method = "satterthwaite", drop = "Intercept")
```


### GLMM:

**On the counts:**

Poisson:

```{r}
Calb_P8_mod_N_CC_poiss <- glmmTMB::glmmTMB(
  N_CC ~ Condition + (1 | Mouse),
  family = poisson("log"),
  data = Calb_P8_data,
  REML = TRUE
)

parameters(Calb_P8_mod_N_CC_poiss)
cat("\n")
performance(Calb_P8_mod_N_CC_poiss)
cat("\n")
check_overdispersion(Calb_P8_mod_N_CC_poiss)
```

Negative binomial:

```{r}
Calb_P8_mod_N_CC_nb <- glmmTMB::glmmTMB(
  N_CC ~ Condition + (1 | Mouse),
  family = nbinom1("log"),
  data = Calb_P8_data,
  REML = TRUE
)

parameters(Calb_P8_mod_N_CC_nb)
cat("\n")
performance(Calb_P8_mod_N_CC_nb)
cat("\n")
check_overdispersion(Calb_P8_mod_N_CC_nb)
```

Generalized Poisson:

```{r}
Calb_P8_mod_N_CC_gp <- glmmTMB::glmmTMB(
  N_CC ~ Condition + (1 | Mouse),
  family = genpois("log"),
  data = Calb_P8_data,
  REML = TRUE
)

parameters(Calb_P8_mod_N_CC_gp)
cat("\n")
performance(Calb_P8_mod_N_CC_gp)
cat("\n")
check_overdispersion(Calb_P8_mod_N_CC_gp)
```

<!------------------------------------------------------------------------------>
## 2. Model Diagnostics:
***

**Model comparison:**

Counts:

```{r}
performance::compare_performance(Calb_P8_mod_N_CC_poiss, Calb_P8_mod_N_CC_nb, Calb_P8_mod_N_CC_gp)
```

Best model:

```{r}
Calb_P8_mod_N_CC <- Calb_P8_mod_N_CC_gp
```


#### Residuals:

```{r fig.width = 8}
performance::check_model(Calb_P8_mod_N_CC)

make_acf_plot(Calb_P8_mod_N_CC)
```


#### Predictions:

```{r}
nsim <- 300

Calb_P8_mod_N_CC_dharma <- DHARMa::simulateResiduals(Calb_P8_mod_N_CC, plot = F, n = nsim, seed = getOption("seed"))
Calb_P8_mod_N_CC_dharma_t <- Calb_P8_mod_N_CC_dharma$simulatedResponse |> t()
```

```{r fig.width = 10}
ppc_plots(Calb_P8_mod_N_CC, simulations = Calb_P8_mod_N_CC_dharma_t, term = "Condition", is_count = FALSE)
```

```{r fig.width = 10}
ppc_stat_plots(Calb_P8_mod_N_CC, simulations = Calb_P8_mod_N_CC_dharma_t, term = "Condition")
```

**Potential outliers:**

```{r fig.width = 10}
insight::get_data(Calb_P8_mod_N_CC) |> 
  rownames_to_column("ID") |> 
  filter(ID %in% DHARMa::outliers(Calb_P8_mod_N_CC_dharma))
```


<!------------------------------------------------------------------------------>
## 3. Effects Analysis:
***

### Coefficients:

**All effects (Wald):**

```{r}
parameters::parameters(
  Calb_P8_mod_N_CC, exponentiate = should_exp(Calb_P8_mod_N_CC),
  ci_method = "Wald", p_adjust = "none", summary = TRUE, digits = 3
)
```

**Main effects (Wald):**

```{r}
car::Anova(Calb_P8_mod_N_CC, type = 3)
```

**Main effects (LRT):**

```{r}
LRT(Calb_P8_mod_N_CC, pred = "Condition")
```


### Marginal Effects:

**Condition:**

Marginal Means:

```{r}
log.main("===[Observed]===\n")

distribution_summary(insight::get_data(Calb_P8_mod_N_CC), dvs = insight::find_response(Calb_P8_mod_N_CC), between = "Condition")

log.main("===[Emmeans]===\n")

emmeans::emmeans(Calb_P8_mod_N_CC, specs = "Condition", type = "response")
```

Marginal Contrasts:

```{r}
log.main("===[Link scale]===\n")

emmeans::emmeans(Calb_P8_mod_N_CC, specs = "Condition", type = "response") |> 
  emmeans::contrast(method = "pairwise", adjust = "none", infer = TRUE)

log.main("===[Response scale]===\n")

emmeans::emmeans(Calb_P8_mod_N_CC, specs = "Condition", trans = "response") |>
  emmeans::contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

Plot:

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Calb_P8_mod_N_CC, xaxis = "Condition", subtitle = stage) |> 
  save_png(filename = glue("[{get_model_tag(Calb_P8_mod_N_CC)}] Boxplots - by [Condition]"), subfolder = Calb_P8_fig_path, dpi = dpi_save_png, width = 5, height = 8)
```



<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
***
# III. Vol_PC_per_cell:
***

<!------------------------------------------------------------------------------>
## 1. Models
***

### Welch t-test:

On data averaged by Mouse.
Data is assumed heteroscedastic in relation to `Condition`.

```{r}
Calb_P8_mod_Vol_PC_per_cell_agg <- afex::mixed(
  Vol_PC_per_cell ~ Condition + (0 + dCondition | ID),
  data = Calb_P8_data_agg,
  control = lmerControl(check.nobs.vs.nlev = "ignore", check.nobs.vs.nRE = "ignore"),
  check_contrasts = FALSE,
  method = "S",
  type = 3,
  test_intercept = FALSE
)

parameters(Calb_P8_mod_Vol_PC_per_cell_agg$full_model, effects = "fixed", ci_method = "satterthwaite", drop = "Intercept")
```


### GLMM:

**Gaussian:**

```{r}
Calb_P8_mod_Vol_PC_per_cell_gauss <- glmmTMB(
  Vol_PC_per_cell ~ Condition + (1 | Mouse),
  family = gaussian("log"),
  data = Calb_P8_data,
  REML = TRUE
)

parameters(Calb_P8_mod_Vol_PC_per_cell_gauss)
cat("\n")
performance(Calb_P8_mod_Vol_PC_per_cell_gauss)
```

**Gamma:**

```{r}
Calb_P8_mod_Vol_PC_per_cell_gamma <- glmmTMB(
  Vol_PC_per_cell ~ Condition + (1 | Mouse),
  family = Gamma("log"),
  data = Calb_P8_data,
  REML = TRUE
)

parameters(Calb_P8_mod_Vol_PC_per_cell_gamma)
cat("\n")
performance(Calb_P8_mod_Vol_PC_per_cell_gamma)
```


<!------------------------------------------------------------------------------>
## 2. Model Diagnostics:
***

**Model comparison:**

```{r}
performance::compare_performance(Calb_P8_mod_Vol_PC_per_cell_gauss, Calb_P8_mod_Vol_PC_per_cell_gamma)
```

Best model:

```{r}
Calb_P8_mod_Vol_PC_per_cell <- Calb_P8_mod_Vol_PC_per_cell_gamma
```


### Residuals:

```{r fig.width = 8}
performance::check_model(Calb_P8_mod_Vol_PC_per_cell)

make_acf_plot(Calb_P8_mod_Vol_PC_per_cell)
```


### Predictions:

```{r}
nsim <- 300

Calb_P8_mod_Vol_PC_per_cell_dharma <- DHARMa::simulateResiduals(Calb_P8_mod_Vol_PC_per_cell, plot = F, n = nsim, seed = getOption("seed"))
Calb_P8_mod_Vol_PC_per_cell_dharma_t <- Calb_P8_mod_Vol_PC_per_cell_dharma$simulatedResponse |> t()
```

```{r fig.width = 10}
ppc_plots(Calb_P8_mod_Vol_PC_per_cell, simulations = Calb_P8_mod_Vol_PC_per_cell_dharma_t, term = "Condition")
```

```{r fig.width = 10}
ppc_stat_plots(Calb_P8_mod_Vol_PC_per_cell, simulations = Calb_P8_mod_Vol_PC_per_cell_dharma_t, term = "Condition")
```

**Potential outliers:**

```{r fig.width = 10}
insight::get_data(Calb_P8_mod_Vol_PC_per_cell) |> 
  rownames_to_column("ID") |> 
  filter(ID %in% DHARMa::outliers(Calb_P8_mod_Vol_PC_per_cell_dharma))
```


<!------------------------------------------------------------------------------>
## 3. Effects Analysis:
***

### Coefficients:

**All effects (Wald):**

```{r}
parameters::parameters(
  Calb_P8_mod_Vol_PC_per_cell, exponentiate = should_exp(Calb_P8_mod_Vol_PC_per_cell),
  ci_method = "Wald", p_adjust = "none", summary = TRUE, digits = 3
)
```

**Main effects (Wald):**

```{r}
car::Anova(Calb_P8_mod_Vol_PC_per_cell, type = 3)
```

**Main effects (LRT):**

```{r}
LRT(Calb_P8_mod_Vol_PC_per_cell, pred = "Condition")
```


### Marginal Effects:

**Condition:**

Marginal Means:

```{r}
log.main("===[Observed]===\n")

distribution_summary(insight::get_data(Calb_P8_mod_Vol_PC_per_cell), dvs = insight::find_response(Calb_P8_mod_Vol_PC_per_cell), between = "Condition")

log.main("===[Emmeans]===\n")

emmeans::emmeans(Calb_P8_mod_Vol_PC_per_cell, specs = "Condition", type = "response")
```

Marginal Contrasts:

```{r}
log.main("===[Link scale]===\n")

emmeans::emmeans(Calb_P8_mod_Vol_PC_per_cell, specs = "Condition", type = "response") |> 
  emmeans::contrast(method = "pairwise", adjust = "none", infer = TRUE)

log.main("===[Response scale]===\n")

emmeans::emmeans(Calb_P8_mod_Vol_PC_per_cell, specs = "Condition", trans = "response") |>
  emmeans::contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

Plot:

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Calb_P8_mod_Vol_PC_per_cell, xaxis = "Condition", subtitle = stage) |> 
  save_png(filename = glue("[{get_model_tag(Calb_P8_mod_Vol_PC_per_cell)}] Boxplots - by [Condition]"), subfolder = Calb_P8_fig_path, dpi = dpi_save_png, width = 5, height = 8)
```

