```{r}
source("../../src/setup.R", echo = FALSE)
```

<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
# I. Data:
***

```{r}
(Calb_P8_data <- load_calb_data() |> map(\(x) filter(x, Stage == "P8")))

Calb_P8_responses <- c("N_CC", "Vol_PC_per_cell")
Calb_P8_predictors <- c("Condition")
```


<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
# II. N_CC
***

<!------------------------------------------------------------------------------>
## 1. Models
***

**Poisson**

```{r}
Calb_P8_mod_N_CC_poiss <- glmmTMB::glmmTMB(
  N_CC ~ Condition + (1 | Mouse),
  family = poisson("log"),
  data = Calb_P8_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Calb_P8_mod_N_CC_poiss)
cli_h1("[Fitness]")
performance(Calb_P8_mod_N_CC_poiss)
cli_h1("[Dispersion]")
check_overdispersion(Calb_P8_mod_N_CC_poiss)
```

**Negative binomial**

```{r}
Calb_P8_mod_N_CC_nb <- glmmTMB::glmmTMB(
  N_CC ~ Condition + (1 | Mouse),
  family = nbinom1("log"),
  data = Calb_P8_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Calb_P8_mod_N_CC_nb)
cli_h1("[Fitness]")
performance(Calb_P8_mod_N_CC_nb)
cli_h1("[Dispersion]")
check_overdispersion(Calb_P8_mod_N_CC_nb)
```

**Generalized Poisson**

```{r}
Calb_P8_mod_N_CC_gp <- glmmTMB::glmmTMB(
  N_CC ~ Condition + (1 | Mouse),
  family = genpois("log"),
  data = Calb_P8_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Calb_P8_mod_N_CC_gp)
cli_h1("[Fitness]")
performance(Calb_P8_mod_N_CC_gp)
cli_h1("[Dispersion]")
check_overdispersion(Calb_P8_mod_N_CC_gp)
```

<!------------------------------------------------------------------------------>
## 2. Model Diagnostics
***

**Model comparison:**

```{r}
compare_performance(Calb_P8_mod_N_CC_poiss, Calb_P8_mod_N_CC_nb, Calb_P8_mod_N_CC_gp)
```

**Most plausible model:**

```{r}
Calb_P8_mod_N_CC <- Calb_P8_mod_N_CC_gp
```


### Residual diagnostics

```{r fig.width = 10}
check_model(Calb_P8_mod_N_CC)

make_acf_plot(Calb_P8_mod_N_CC)
```


### Predictive checks

```{r}
Calb_P8_mod_N_CC_dharma <- DHARMa::simulateResiduals(Calb_P8_mod_N_CC, plot = FALSE, n = 300, seed = getOption("seed"))
Calb_P8_mod_N_CC_dharma_t <- t(Calb_P8_mod_N_CC_dharma$simulatedResponse)
```

```{r fig.width = 10}
ppc_plots(Calb_P8_mod_N_CC, simulations = Calb_P8_mod_N_CC_dharma_t, term = "Condition", is_count = FALSE)
```

```{r fig.width = 10}
ppc_stat_plots(Calb_P8_mod_N_CC, simulations = Calb_P8_mod_N_CC_dharma_t, term = "Condition")
```

### Potential outliers

```{r}
get_model_based_outliers(Calb_P8_data$clean, Calb_P8_mod_N_CC, Calb_P8_mod_N_CC_dharma, Calb_P8_responses)
```


<!------------------------------------------------------------------------------>
## 3. Model Analysis
***

### Model parameters

**All effects (Wald):**

```{r}
parameters(
  Calb_P8_mod_N_CC, exponentiate = should_exp(Calb_P8_mod_N_CC),
  ci_method = "Wald", p_adjust = "none", summary = TRUE, digits = 3
)
```

**Main effects (Wald):**

```{r}
car::Anova(Calb_P8_mod_N_CC, type = 3)
```

### Marginal means

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Calb_P8_mod_N_CC), dvs = find_response(Calb_P8_mod_N_CC), between = "Condition")

cli_h1("[Emmeans]")

emmeans(Calb_P8_mod_N_CC, specs = "Condition", type = "response")
```

### Contrasts

```{r}
cli_h1("[Link scale]")

emmeans(Calb_P8_mod_N_CC, specs = "Condition", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Calb_P8_mod_N_CC, specs = "Condition", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

Plot:

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Calb_P8_mod_N_CC, xaxis = "Condition", subtitle = unique(Calb_P8_data$clean$Stage))
```



<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
# III. Vol_PC_per_cell
***

<!------------------------------------------------------------------------------>
## 1. Models
***

**Gaussian:**

```{r}
Calb_P8_mod_Vol_PC_gauss <- glmmTMB(
  Vol_PC_per_cell ~ Condition + (1 | Mouse),
  family = gaussian("log"),
  data = Calb_P8_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Calb_P8_mod_Vol_PC_gauss)
cli_h1("[Fitness]")
performance(Calb_P8_mod_Vol_PC_gauss)
```

**Gamma:**

```{r}
Calb_P8_mod_Vol_PC_gamma <- glmmTMB(
  Vol_PC_per_cell ~ Condition + (1 | Mouse),
  family = Gamma("log"),
  data = Calb_P8_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Calb_P8_mod_Vol_PC_gamma)
cli_h1("[Fitness]")
performance(Calb_P8_mod_Vol_PC_gamma)
```


<!------------------------------------------------------------------------------>
## 2. Model Diagnostics
***

**Model comparison:**

```{r}
compare_performance(Calb_P8_mod_Vol_PC_gauss, Calb_P8_mod_Vol_PC_gamma)
```

Best model:

```{r}
Calb_P8_mod_Vol_PC <- Calb_P8_mod_Vol_PC_gamma
```


### Residual diagnostics

```{r fig.width = 10}
check_model(Calb_P8_mod_Vol_PC)

make_acf_plot(Calb_P8_mod_Vol_PC)
```


### Predictive checks

```{r}
Calb_P8_mod_Vol_PC_dharma <- DHARMa::simulateResiduals(Calb_P8_mod_Vol_PC, plot = FALSE, n = 300, seed = getOption("seed"))
Calb_P8_mod_Vol_PC_dharma_t <- t(Calb_P8_mod_Vol_PC_dharma$simulatedResponse)
```

```{r fig.width = 10}
ppc_plots(Calb_P8_mod_Vol_PC, simulations = Calb_P8_mod_Vol_PC_dharma_t, term = "Condition")
```

```{r fig.width = 10}
ppc_stat_plots(Calb_P8_mod_Vol_PC, simulations = Calb_P8_mod_Vol_PC_dharma_t, term = "Condition")
```

### Potential outliers

```{r}
get_model_based_outliers(Calb_P8_data$clean, Calb_P8_mod_Vol_PC, Calb_P8_mod_Vol_PC_dharma, Calb_P8_responses)
```


<!------------------------------------------------------------------------------>
## 3. Model Analysis
***

### Model parameters

**All effects (Wald):**

```{r}
parameters(
  Calb_P8_mod_Vol_PC, exponentiate = should_exp(Calb_P8_mod_Vol_PC),
  ci_method = "Wald", p_adjust = "none", summary = TRUE, digits = 3
)
```

**Main effects (Wald):**

```{r}
car::Anova(Calb_P8_mod_Vol_PC, type = 3)
```


### Marginal means

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Calb_P8_mod_Vol_PC), dvs = find_response(Calb_P8_mod_Vol_PC), between = "Condition")

cli_h1("[Emmeans]")

emmeans(Calb_P8_mod_Vol_PC, specs = "Condition", type = "response")
```


### Contrasts

```{r}
cli_h1("[Link scale]")

emmeans(Calb_P8_mod_Vol_PC, specs = "Condition", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Calb_P8_mod_Vol_PC, specs = "Condition", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```


Plot:

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Calb_P8_mod_Vol_PC, xaxis = "Condition", subtitle = unique(Calb_P8_data$clean$Stage))
```

