```{r}
source("../../src/setup.R", echo = FALSE)
```

<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
# I. Data
***

```{r}
(Casp_P4_data <- load_casp_data() |> map(\(x) filter(x, Stage == "P4")))

Casp_P4_responses <- c("Dens_Tot", "A_EGL", "A_ML_PCL", "A_IGL", "A_WM")

Casp_P4_predictors <- c("Condition", "Z")
```


<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
# II. Dens_Tot
***

<!------------------------------------------------------------------------------>
## 1. Models
***

**Gaussian:**

```{r}
Casp_P4_mod_Dens_Tot_gauss <- glmmTMB(
  Dens_Tot ~ Condition * Z + (1 | Mouse),
  family = gaussian("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_Dens_Tot_gauss)
cli_h1("[Fitness]")
performance(Casp_P4_mod_Dens_Tot_gauss)
```

**Gamma:**

```{r}
Casp_P4_mod_Dens_Tot_gamma <- glmmTMB(
  Dens_Tot ~ Condition * Z + (1 | Mouse),
  family = Gamma("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_Dens_Tot_gamma)
cli_h1("[Fitness]")
performance(Casp_P4_mod_Dens_Tot_gamma)
```


<!------------------------------------------------------------------------------>
## 2. Model Diagnostics
***

**Model comparison:**

```{r}
compare_performance(Casp_P4_mod_Dens_Tot_gauss, Casp_P4_mod_Dens_Tot_gamma)
```

Best model:

```{r}
Casp_P4_mod_Dens_Tot <- Casp_P4_mod_Dens_Tot_gamma
```


### Residual diagnostics

```{r fig.width = 10}
check_model(Casp_P4_mod_Dens_Tot)

make_acf_plot(Casp_P4_mod_Dens_Tot)
```


### Predictive checks

```{r}
Casp_P4_mod_Dens_Tot_dharma <- DHARMa::simulateResiduals(Casp_P4_mod_Dens_Tot, plot = FALSE, n = 300, seed = getOption("seed"))
Casp_P4_mod_Dens_Tot_dharma_t <- t(Casp_P4_mod_Dens_Tot_dharma$simulatedResponse)
```

```{r fig.width = 10}
ppc_plots(Casp_P4_mod_Dens_Tot, simulations = Casp_P4_mod_Dens_Tot_dharma_t, term = "Condition")
ppc_plots(Casp_P4_mod_Dens_Tot, simulations = Casp_P4_mod_Dens_Tot_dharma_t, term = "Z")
```

```{r fig.width = 10}
ppc_stat_plots(Casp_P4_mod_Dens_Tot, simulations = Casp_P4_mod_Dens_Tot_dharma_t, term = "Condition")
ppc_stat_plots(Casp_P4_mod_Dens_Tot, simulations = Casp_P4_mod_Dens_Tot_dharma_t, term = "Z")
```


### Potential outliers

```{r}
get_model_based_outliers(Casp_P4_data$clean, Casp_P4_mod_Dens_Tot, Casp_P4_mod_Dens_Tot_dharma, Casp_P4_responses)
```


<!------------------------------------------------------------------------------>
## 3. Model Analysis
***

### Model parameters

**All effects (Wald):**

```{r}
parameters(
  Casp_P4_mod_Dens_Tot, exponentiate = should_exp(Casp_P4_mod_Dens_Tot),
  ci_method = "Wald", p_adjust = "none", summary = TRUE, digits = 3
)
```

**Main effects (Wald):**

```{r}
car::Anova(Casp_P4_mod_Dens_Tot, type = 3)
```


### Marginal means

**Condition**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_Dens_Tot), dvs = find_response(Casp_P4_mod_Dens_Tot), between = "Condition")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_Dens_Tot, specs = "Condition", type = "response")
```

**Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_Dens_Tot), dvs = find_response(Casp_P4_mod_Dens_Tot), between = "Z")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_Dens_Tot, specs = "Z", type = "response")
```

**Condition:Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_Dens_Tot), dvs = find_response(Casp_P4_mod_Dens_Tot), between = c("Condition", "Z"))

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_Dens_Tot, specs = ~ Condition | Z, type = "response")
```


### Contrasts

**Condition**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_Dens_Tot, specs = "Condition", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_Dens_Tot, specs = "Condition", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_Dens_Tot, xaxis = "Condition", subtitle = unique(Casp_P4_data$clean$Stage))
```

**Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_Dens_Tot, specs = "Z", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_Dens_Tot, specs = "Z", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_Dens_Tot, xaxis = "Z")
```

**Condition:Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_Dens_Tot, specs = ~ Condition | Z, type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_Dens_Tot, specs = ~ Condition | Z, type = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_Dens_Tot, specs = ~ Condition | Z, regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_Dens_Tot, specs = ~ Condition | Z, regrid = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot(Casp_P4_mod_Dens_Tot, xaxis = "Condition", facet = "Z")
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot_inter(Casp_P4_mod_Dens_Tot, pred1 = "Condition", pred2 = "Z")
```



<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
# III. A_EGL
***

<!------------------------------------------------------------------------------>
## 1. Models
***

**Gaussian:**

```{r}
Casp_P4_mod_A_EGL_gauss <- glmmTMB(
  A_EGL ~ Condition * Z + offset(log(A_Tot)) + (1 | Mouse),
  family = gaussian("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_A_EGL_gauss)
cli_h1("[Fitness]")
performance(Casp_P4_mod_A_EGL_gauss)
```

**Gamma:**

```{r}
Casp_P4_mod_A_EGL_gamma <- glmmTMB(
  A_EGL ~ Condition * Z + offset(log(A_Tot)) + (1 | Mouse),
  family = Gamma("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_A_EGL_gamma)
cli_h1("[Fitness]")
performance(Casp_P4_mod_A_EGL_gamma)
```


<!------------------------------------------------------------------------------>
## 2. Model Diagnostics
***

**Model comparison:**

```{r}
compare_performance(Casp_P4_mod_A_EGL_gauss, Casp_P4_mod_A_EGL_gamma)
```

Best model:

```{r}
Casp_P4_mod_A_EGL <- Casp_P4_mod_A_EGL_gamma
```


### Residual diagnostics

```{r fig.width = 10}
check_model(Casp_P4_mod_A_EGL)

make_acf_plot(Casp_P4_mod_A_EGL)
```


### Predictive checks

```{r}
Casp_P4_mod_A_EGL_dharma <- DHARMa::simulateResiduals(Casp_P4_mod_A_EGL, plot = FALSE, n = 300, seed = getOption("seed"))
Casp_P4_mod_A_EGL_dharma_t <- t(Casp_P4_mod_A_EGL_dharma$simulatedResponse)
```

```{r fig.width = 10}
ppc_plots(Casp_P4_mod_A_EGL, simulations = Casp_P4_mod_A_EGL_dharma_t, term = "Condition")
ppc_plots(Casp_P4_mod_A_EGL, simulations = Casp_P4_mod_A_EGL_dharma_t, term = "Z")
```

```{r fig.width = 10}
ppc_stat_plots(Casp_P4_mod_A_EGL, simulations = Casp_P4_mod_A_EGL_dharma_t, term = "Condition")
ppc_stat_plots(Casp_P4_mod_A_EGL, simulations = Casp_P4_mod_A_EGL_dharma_t, term = "Z")
```


### Potential outliers

```{r}
get_model_based_outliers(Casp_P4_data$clean, Casp_P4_mod_A_EGL, Casp_P4_mod_A_EGL_dharma, Casp_P4_responses)
```


<!------------------------------------------------------------------------------>
## 3. Model Analysis
***

### Model parameters

**All effects (Wald):**

```{r}
parameters(
  Casp_P4_mod_A_EGL, exponentiate = should_exp(Casp_P4_mod_A_EGL),
  ci_method = "Wald", p_adjust = "none", summary = TRUE, digits = 3
)
```

**Main effects (Wald):**

```{r}
car::Anova(Casp_P4_mod_A_EGL, type = 3)
```


### Marginal means

**Condition**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_EGL), dvs = find_response(Casp_P4_mod_A_EGL), between = "Condition")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_EGL, specs = "Condition", type = "response")
```

**Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_EGL), dvs = find_response(Casp_P4_mod_A_EGL), between = "Z")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_EGL, specs = "Z", type = "response")
```

**Condition:Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_EGL), dvs = find_response(Casp_P4_mod_A_EGL), between = c("Condition", "Z"))

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_EGL, specs = ~ Condition | Z, type = "response")
```


### Contrasts

**Condition**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_EGL, specs = "Condition", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_EGL, specs = "Condition", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_A_EGL, xaxis = "Condition", subtitle = unique(Casp_P4_data$clean$Stage))
```

**Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_EGL, specs = "Z", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_EGL, specs = "Z", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_A_EGL, xaxis = "Z")
```

**Condition:Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_EGL, specs = ~ Condition | Z, type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_A_EGL, specs = ~ Condition | Z, type = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_EGL, specs = ~ Condition | Z, regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_A_EGL, specs = ~ Condition | Z, regrid = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot(Casp_P4_mod_A_EGL, xaxis = "Condition", facet = "Z")
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot_inter(Casp_P4_mod_A_EGL, pred1 = "Condition", pred2 = "Z")
```

<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
# IV. A_ML_PCL
***

<!------------------------------------------------------------------------------>
## 1. Models
***

**Gaussian:**

```{r}
Casp_P4_mod_A_ML_PCL_gauss <- glmmTMB(
  A_ML_PCL ~ Condition * Z + offset(log(A_Tot)) + (1 | Mouse),
  family = gaussian("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_A_ML_PCL_gauss)
cli_h1("[Fitness]")
performance(Casp_P4_mod_A_ML_PCL_gauss)
```

**Gamma:**

```{r}
Casp_P4_mod_A_ML_PCL_gamma <- glmmTMB(
  A_ML_PCL ~ Condition * Z + offset(log(A_Tot)) + (1 | Mouse),
  family = Gamma("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_A_ML_PCL_gamma)
cli_h1("[Fitness]")
performance(Casp_P4_mod_A_ML_PCL_gamma)
```


<!------------------------------------------------------------------------------>
## 2. Model Diagnostics
***

**Model comparison:**

```{r}
compare_performance(Casp_P4_mod_A_ML_PCL_gauss, Casp_P4_mod_A_ML_PCL_gamma)
```

Best model:

```{r}
Casp_P4_mod_A_ML_PCL <- Casp_P4_mod_A_ML_PCL_gamma
```


### Residual diagnostics

```{r fig.width = 10}
check_model(Casp_P4_mod_A_ML_PCL)

make_acf_plot(Casp_P4_mod_A_ML_PCL)
```


### Predictive checks

```{r}
Casp_P4_mod_A_ML_PCL_dharma <- DHARMa::simulateResiduals(Casp_P4_mod_A_ML_PCL, plot = FALSE, n = 300, seed = getOption("seed"))
Casp_P4_mod_A_ML_PCL_dharma_t <- t(Casp_P4_mod_A_ML_PCL_dharma$simulatedResponse)
```

```{r fig.width = 10}
ppc_plots(Casp_P4_mod_A_ML_PCL, simulations = Casp_P4_mod_A_ML_PCL_dharma_t, term = "Condition")
ppc_plots(Casp_P4_mod_A_ML_PCL, simulations = Casp_P4_mod_A_ML_PCL_dharma_t, term = "Z")
```

```{r fig.width = 10}
ppc_stat_plots(Casp_P4_mod_A_ML_PCL, simulations = Casp_P4_mod_A_ML_PCL_dharma_t, term = "Condition")
ppc_stat_plots(Casp_P4_mod_A_ML_PCL, simulations = Casp_P4_mod_A_ML_PCL_dharma_t, term = "Z")
```


### Potential outliers

```{r}
get_model_based_outliers(Casp_P4_data$clean, Casp_P4_mod_A_ML_PCL, Casp_P4_mod_A_ML_PCL_dharma, Casp_P4_responses)
```


<!------------------------------------------------------------------------------>
## 3. Model Analysis
***

### Model parameters

**All effects (Wald):**

```{r}
parameters(
  Casp_P4_mod_A_ML_PCL, exponentiate = should_exp(Casp_P4_mod_A_ML_PCL),
  ci_method = "Wald", p_adjust = "none", summary = TRUE, digits = 3
)
```

**Main effects (Wald):**

```{r}
car::Anova(Casp_P4_mod_A_ML_PCL, type = 3)
```


### Marginal means

**Condition**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_ML_PCL), dvs = find_response(Casp_P4_mod_A_ML_PCL), between = "Condition")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = "Condition", type = "response")
```

**Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_ML_PCL), dvs = find_response(Casp_P4_mod_A_ML_PCL), between = "Z")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = "Z", type = "response")
```

**Condition:Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_ML_PCL), dvs = find_response(Casp_P4_mod_A_ML_PCL), between = c("Condition", "Z"))

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = ~ Condition | Z, type = "response")
```


### Contrasts

**Condition**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = "Condition", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = "Condition", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_A_ML_PCL, xaxis = "Condition", subtitle = unique(Casp_P4_data$clean$Stage))
```

**Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = "Z", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = "Z", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_A_ML_PCL, xaxis = "Z")
```

**Condition:Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = ~ Condition | Z, type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = ~ Condition | Z, type = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = ~ Condition | Z, regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_A_ML_PCL, specs = ~ Condition | Z, regrid = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot(Casp_P4_mod_A_ML_PCL, xaxis = "Condition", facet = "Z")
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot_inter(Casp_P4_mod_A_ML_PCL, pred1 = "Condition", pred2 = "Z")
```

<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
# V. A_IGL
***

<!------------------------------------------------------------------------------>
## 1. Models
***

**Gaussian:**

```{r}
Casp_P4_mod_A_IGL_gauss <- glmmTMB(
  A_IGL ~ Condition * Z + offset(log(A_Tot)) + (1 | Mouse),
  family = gaussian("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_A_IGL_gauss)
cli_h1("[Fitness]")
performance(Casp_P4_mod_A_IGL_gauss)
```

**Gamma:**

```{r}
Casp_P4_mod_A_IGL_gamma <- glmmTMB(
  A_IGL ~ Condition * Z + offset(log(A_Tot)) + (1 | Mouse),
  family = Gamma("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_A_IGL_gamma)
cli_h1("[Fitness]")
performance(Casp_P4_mod_A_IGL_gamma)
```


<!------------------------------------------------------------------------------>
## 2. Model Diagnostics
***

**Model comparison:**

```{r}
compare_performance(Casp_P4_mod_A_IGL_gauss, Casp_P4_mod_A_IGL_gamma)
```

Best model:

```{r}
Casp_P4_mod_A_IGL <- Casp_P4_mod_A_IGL_gamma
```


### Residual diagnostics

```{r fig.width = 10}
check_model(Casp_P4_mod_A_IGL)

make_acf_plot(Casp_P4_mod_A_IGL)
```


### Predictive checks

```{r}
Casp_P4_mod_A_IGL_dharma <- DHARMa::simulateResiduals(Casp_P4_mod_A_IGL, plot = FALSE, n = 300, seed = getOption("seed"))
Casp_P4_mod_A_IGL_dharma_t <- t(Casp_P4_mod_A_IGL_dharma$simulatedResponse)
```

```{r fig.width = 10}
ppc_plots(Casp_P4_mod_A_IGL, simulations = Casp_P4_mod_A_IGL_dharma_t, term = "Condition")
ppc_plots(Casp_P4_mod_A_IGL, simulations = Casp_P4_mod_A_IGL_dharma_t, term = "Z")
```

```{r fig.width = 10}
ppc_stat_plots(Casp_P4_mod_A_IGL, simulations = Casp_P4_mod_A_IGL_dharma_t, term = "Condition")
ppc_stat_plots(Casp_P4_mod_A_IGL, simulations = Casp_P4_mod_A_IGL_dharma_t, term = "Z")
```


### Potential outliers

```{r}
get_model_based_outliers(Casp_P4_data$clean, Casp_P4_mod_A_IGL, Casp_P4_mod_A_IGL_dharma, Casp_P4_responses)
```


<!------------------------------------------------------------------------------>
## 3. Model Analysis
***

### Model parameters

**All effects (Wald):**

```{r}
parameters(
  Casp_P4_mod_A_IGL, exponentiate = should_exp(Casp_P4_mod_A_IGL),
  ci_method = "Wald", p_adjust = "none", summary = TRUE, digits = 3
)
```

**Main effects (Wald):**

```{r}
car::Anova(Casp_P4_mod_A_IGL, type = 3)
```


### Marginal means

**Condition**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_IGL), dvs = find_response(Casp_P4_mod_A_IGL), between = "Condition")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_IGL, specs = "Condition", type = "response")
```

**Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_IGL), dvs = find_response(Casp_P4_mod_A_IGL), between = "Z")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_IGL, specs = "Z", type = "response")
```

**Condition:Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_IGL), dvs = find_response(Casp_P4_mod_A_IGL), between = c("Condition", "Z"))

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_IGL, specs = ~ Condition | Z, type = "response")
```


### Contrasts

**Condition**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_IGL, specs = "Condition", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_IGL, specs = "Condition", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_A_IGL, xaxis = "Condition", subtitle = unique(Casp_P4_data$clean$Stage))
```

**Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_IGL, specs = "Z", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_IGL, specs = "Z", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_A_IGL, xaxis = "Z")
```

**Condition:Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_IGL, specs = ~ Condition | Z, type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_A_IGL, specs = ~ Condition | Z, type = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_IGL, specs = ~ Condition | Z, regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_A_IGL, specs = ~ Condition | Z, regrid = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot(Casp_P4_mod_A_IGL, xaxis = "Condition", facet = "Z")
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot_inter(Casp_P4_mod_A_IGL, pred1 = "Condition", pred2 = "Z")
```



<!------------------------------------------------------------------------------>
<!------------------------------------------------------------------------------>
# VI. A_WM
***

<!------------------------------------------------------------------------------>
## 1. Models
***

**Gaussian:**

```{r}
Casp_P4_mod_A_WM_gauss <- glmmTMB(
  A_WM ~ Condition * Z + offset(log(A_Tot)) + (1 | Mouse),
  family = gaussian("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_A_WM_gauss)
cli_h1("[Fitness]")
performance(Casp_P4_mod_A_WM_gauss)
```

**Gamma:**

```{r}
Casp_P4_mod_A_WM_gamma <- glmmTMB(
  A_WM ~ Condition * Z + offset(log(A_Tot)) + (1 | Mouse),
  family = Gamma("log"),
  data = Casp_P4_data$clean,
  REML = TRUE
)

cli_h1("[Parameters]")
parameters(Casp_P4_mod_A_WM_gamma)
cli_h1("[Fitness]")
performance(Casp_P4_mod_A_WM_gamma)
```


<!------------------------------------------------------------------------------>
## 2. Model Diagnostics
***

**Model comparison:**

```{r}
compare_performance(Casp_P4_mod_A_WM_gauss, Casp_P4_mod_A_WM_gamma)
```

Best model:

```{r}
Casp_P4_mod_A_WM <- Casp_P4_mod_A_WM_gamma
```


### Residual diagnostics

```{r fig.width = 10}
check_model(Casp_P4_mod_A_WM)

make_acf_plot(Casp_P4_mod_A_WM)
```


### Predictive checks

```{r}
Casp_P4_mod_A_WM_dharma <- DHARMa::simulateResiduals(Casp_P4_mod_A_WM, plot = FALSE, n = 300, seed = getOption("seed"))
Casp_P4_mod_A_WM_dharma_t <- t(Casp_P4_mod_A_WM_dharma$simulatedResponse)
```

```{r fig.width = 10}
ppc_plots(Casp_P4_mod_A_WM, simulations = Casp_P4_mod_A_WM_dharma_t, term = "Condition")
ppc_plots(Casp_P4_mod_A_WM, simulations = Casp_P4_mod_A_WM_dharma_t, term = "Z")
```

```{r fig.width = 10}
ppc_stat_plots(Casp_P4_mod_A_WM, simulations = Casp_P4_mod_A_WM_dharma_t, term = "Condition")
ppc_stat_plots(Casp_P4_mod_A_WM, simulations = Casp_P4_mod_A_WM_dharma_t, term = "Z")
```


### Potential outliers

```{r}
get_model_based_outliers(Casp_P4_data$clean, Casp_P4_mod_A_WM, Casp_P4_mod_A_WM_dharma, Casp_P4_responses)
```


<!------------------------------------------------------------------------------>
## 3. Model Analysis
***

### Model parameters

**All effects (Wald):**

```{r}
parameters(
  Casp_P4_mod_A_WM, exponentiate = should_exp(Casp_P4_mod_A_WM),
  ci_method = "Wald", p_adjust = "none", summary = TRUE, digits = 3
)
```

**Main effects (Wald):**

```{r}
car::Anova(Casp_P4_mod_A_WM, type = 3)
```


### Marginal means

**Condition**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_WM), dvs = find_response(Casp_P4_mod_A_WM), between = "Condition")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_WM, specs = "Condition", type = "response")
```

**Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_WM), dvs = find_response(Casp_P4_mod_A_WM), between = "Z")

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_WM, specs = "Z", type = "response")
```

**Condition:Z**

```{r}
cli_h1("[Observed]")

distribution_summary(get_data(Casp_P4_mod_A_WM), dvs = find_response(Casp_P4_mod_A_WM), between = c("Condition", "Z"))

cli_h1("[Emmeans]")

emmeans(Casp_P4_mod_A_WM, specs = ~ Condition | Z, type = "response")
```


### Contrasts

**Condition**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_WM, specs = "Condition", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_WM, specs = "Condition", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_A_WM, xaxis = "Condition", subtitle = unique(Casp_P4_data$clean$Stage))
```

**Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_WM, specs = "Z", type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_WM, specs = "Z", regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)
```

```{r fig.width = 4, fig.height = 7}
make_signif_boxplot(Casp_P4_mod_A_WM, xaxis = "Z")
```

**Condition:Z**

```{r}
cli_h1("[Link scale]")

emmeans(Casp_P4_mod_A_WM, specs = ~ Condition | Z, type = "response") |> 
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_A_WM, specs = ~ Condition | Z, type = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)

cli_h1("[Response scale]")

emmeans(Casp_P4_mod_A_WM, specs = ~ Condition | Z, regrid = "response") |>
  contrast(method = "pairwise", adjust = "none", infer = TRUE)

cli_h2("[Interaction]")

emmeans(Casp_P4_mod_A_WM, specs = ~ Condition | Z, regrid = "response") |>
  contrast(interaction = "pairwise", by = NULL, adjust = "none", infer = TRUE)
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot(Casp_P4_mod_A_WM, xaxis = "Condition", facet = "Z")
```

```{r fig.width = 6, fig.height = 5}
make_signif_boxplot_inter(Casp_P4_mod_A_WM, pred1 = "Condition", pred2 = "Z")
```
